---
title: "Class 10: Halloween Mini-Project"
author: "Andrew Nguyen"
format: html
---

```{r}
candy_file <- "candy-data.csv"
candy <- read.csv(candy_file, row.names=1)
head(candy)
```

```{r}
nrow(candy)
nrow("fruity")
```

> Q1. How many different candy types are in this dataset?

There are 85 different candy types in this dataset


```{r}
sum(candy$fruity, na.rm= TRUE)
```


> Q2. How many fruity candy types are in the dataset?

There are 38 fruity types in the dataset

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Sour Patch Kids",]$winpercent
```
The winpercent value for Sour Patch Kids is 59.86

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```
The winpercent for Kit Kat is 76.77

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```
The winpercent for Tootsie Roll Snack Bars is 49.63

```{r}
#install.packages("skimr")
```

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

One that seems to be different compared to the majority of the other columns is p100, since we see that most of the values are 1 or near the value 1

> Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}
candy$chocolate
```
I think the zeros and ones represent whether the candy has chocholate (1) or no chocolate (0).

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)

```
> Q9. Is the distribution of winpercent values symmetrical?

The distribution fo winpercent is not symmetrical

> Q10. Is the center of the distribution above or below 50%?

It's below 50%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
choc_per <- candy$winpercent[as.logical(candy$chocolate)]
choc_per

```
```{r}
mean(choc_per)
```


```{r}
fruit_per <- candy$winpercent[as.logical(candy$fruity)]
fruit_per
```
```{r}
mean(fruit_per)
```


Chocolate candy is higher ranked than fruit candy

> Q12. Is this difference statistically significant?

```{r}
t.test(choc_per, fruit_per)
```

The difference between the two are statistically significant since the p=value is lower than 0.05 and the p value is 2.871e-08


> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```
The least liked candy are Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(-candy$winpercent),], n=5)
```

The Top 5 favorite candy of all time in this set are Reese's Peanut Butter Cup, Reeses's Minatures, Twix, Kit Kat, and Snickers

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library("ggplot2")
```

```{r}
ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```



```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

> Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Sixlets

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is Starburst


```{r}
#install.packages("ggrepel")
```

```{r}
library("ggrepel")
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

The highest ranked is Reeses Minatures since it has an over 80 winpercent while also having a low pricepercent of near 0.25

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

The 5 most expensive candy types in the dataset are Nik L Nip, Nestle Smarties, Ring Pop, Hershey's Krackle, and Hershey's Milk Chocolate


> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
#install.packages("corrplot")
```

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Two variables that are anti-correlated are fruity and chocolate

> Q23. Similarly, what two variables are most positively correlated?

the most positively correlated variables are chocolate and bar for the type of candy, but the highest correlated in general seems to be chocolate and winpercent



```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16 )
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#install.packages("plotly")
```

```{r}
library(plotly)
ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Original variables that are picked up strongly by PC1 are fruity, hard, and pluribus.